Estimating Sleep Quality Metrics Using Free-Living Accelerometer Data From Thigh-Worn Devices in Comparison to an EEG-Based Sleep Tracking Device

### 2023-06-27 manuscript/manuscript.tex
pr and roc curves

\usepackage{lineno}\linenumbers
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{npj Digital Medicine}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Estimating Sleep Quality Metrics Using Free-Living Accelerometer Data From Thigh-Worn Devices in Comparison to an EEG-Based Sleep Tracking Device},
  pdfauthor={Esben Høegholm Lykke; Jan Christian Brønd},
  pdfkeywords={Sleep, Accelerometry, EEG, Machine learning, Sleep
quality metrics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{keyword}
    Sleep \sep Accelerometry \sep EEG \sep Machine learning \sep 
    Sleep quality metrics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden, sharp corners, frame hidden, breakable, enhanced]}{\end{tcolorbox}}\fi

@@

\hypertarget{tbl-summary_zm}{}
\begin{longtable}[]{@{}llllllll@{}}
\caption{\label{tbl-summary_zm}Summary of sleep quality metrics as
calculated based on the raw, the 5-minute, and 10-minute median filtered
raw ZM prediction.}\tabularnewline
\toprule\noalign{}
& & Raw & & Median 5-Minute & & Median 10-Minute & \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
& & Raw & & Median 5-Minute & & Median 10-Minute & \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
& & Mean & SD & Mean & SD & Mean & SD \\
& SPT & 9.2 & 2.1 & 9.2 & 2.1 & 9.2 & 2.1 \\
& TST & 7.7 & 1.9 & 8.1 & 2.0 & 8.2 & 2.1 \\
& SE & 82.6 & 12.0 & 86.4 & 12.7 & 87.5 & 12.9 \\
& LPS & 34.5 & 27.9 & 36.3 & 39.8 & 38.0 & 48.7 \\
& WASO & 39.0 & 33.6 & 30.6 & 46.8 & 22.3 & 55.4 \\
\end{longtable}

>>>>

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{lineno}\linenumbers
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{npj Digital Medicine}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Estimating Sleep Quality Metrics Using Free-Living Accelerometer Data From Thigh-Worn Devices in Comparison to an EEG-Based Sleep Tracking Device},
  pdfauthor={Esben Høegholm Lykke; Jan Christian Brønd},
  pdfkeywords={Sleep, Accelerometry, EEG, Machine learning, Sleep
quality metrics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{keyword}
    Sleep \sep Accelerometry \sep EEG \sep Machine learning \sep 
    Sleep quality metrics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi

@@

\hypertarget{tbl-summary_zm}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2400}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1500}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1500}}@{}}
\caption{\label{tbl-summary_zm}Summary of Sleep Quality Metrics based on
Raw ZM Predictions and Median Filtering.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
SPT
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
TST
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
SE
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
LPS
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
WASO
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
SPT
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
TST
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
SE
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
LPS
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
WASO
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Raw & 9.2 (2.1) & 7.7 (1.9) & 82.6 (12.0) & 34.5 (27.9) & 39.0 (33.6) \\
5-Min Median & 9.2 (2.1) & 8.1 (2.0) & 86.4 (12.7) & 36.3 (39.8) & 30.6
(46.8) \\
10-Min Median & 9.2 (2.1) & 8.2 (2.1) & 87.5 (12.9) & 38.0 (48.7) & 22.3
(55.4) \\
\end{longtable}

\hypertarget{performance-on-epoch-to-epoch-basis}{%
\subsection{Performance on Epoch-To-Epoch
Basis}\label{performance-on-epoch-to-epoch-basis}}

The results from the epoch-to-epoch evaluation of in-bed performance
metrics for our included models are presented in
Table~\ref{tbl-in_bed_performance}. The epoch-to-epoch evaluation of
in-bed performance metrics, outlined in @tbl-in\_bed\_performance,
demonstrates practically equivalent performance across all model types.
While the Decision Tree model posted an F1 score of 94.4\% and accuracy
of 95.3\%, the Logistic Regression and Feed-Forward Neural Network
models each exhibited an F1 score of 95.0\% and similar accuracies of
95.7\% and 95.8\%, respectively. The XGBoost model, despite recording
the highest metrics with an F1 score of 95.4\% and accuracy of 96.1\%,
outpaced the others only marginally. This underscores the consistency of
performance among these models in monitoring in-bed conditions.

\hypertarget{tbl-in_bed_performance}{}
\begin{longtable}{lrrrrr}
\caption{\label{tbl-in_bed_performance}In-Bed Performance Metrics }\tabularnewline

\toprule
model & F1 Score (\%) & Accuracy (\%) & Sensitivity (\%) & Precision (\%) & Specificity (\%) \\ 
\midrule
Decision Tree & $9,436.3\%$ & $9,527.2\%$ & $9,312.3\%$ & $9,563.6\%$ & $9,686.0\%$ \\ 
Logistic Regression & $9,498.9\%$ & $9,573.9\%$ & $9,503.5\%$ & $9,494.3\%$ & $9,626.0\%$ \\ 
Feed-Forward Neural Network & $9,502.9\%$ & $9,577.4\%$ & $9,507.1\%$ & $9,498.7\%$ & $9,629.2\%$ \\ 
XGBoost & $9,538.4\%$ & $9,605.8\%$ & $9,582.7\%$ & $9,494.4\%$ & $9,622.9\%$ \\ 
\bottomrule
\end{longtable}

Table~\ref{tbl-sleep_performance} details the performance metrics of the
four machine learning models---Decision Tree, Logistic Regression,
Neural Network, and XGBoost---applied to raw ZM predictions and 5- and
10-minute median filtered predictions. For raw ZM predictions, all
models demonstrated comparable performance, with F1 Scores between
93.27\% and 93.58\%. The XGBoost model performed marginally better.
However, it's crucial to highlight that Specificity was notably low
across models, with the Neural Network model achieving the highest at
48.31\%. Performance improved on 5-minute median data. XGBoost again
outperformed other models in F1 Score and Accuracy, yet Specificity
remained a concern, with scores ranging from 37.47\% to 50.90\%. On
10-minute median data, all models showed further improvement, and
XGBoost continued to lead in F1 Score and Accuracy. It is significant
that both Neural Network and XGBoost models achieved a Sensitivity of
99.10\%. However, low Specificity scores persisted, with the highest
only at 50.98\% for the XGBoost model.

\hypertarget{tbl-sleep_performance}{}
\begin{longtable}{cccccc}
\caption{\label{tbl-sleep_performance}Sleep Performance Metrics }\tabularnewline

\toprule
 & F1 Score (\%) & Accuracy (\%) & Sensitivity (\%) & Precision (\%) & Specificity (\%) \\ 
\midrule
\multicolumn{6}{l}{Raw ZM Predictions} \\ 
\midrule
Decision Tree & $93.27$ & $88.25$ & $96.92$ & $89.88$ & $42.73$ \\ 
Logistic Regression & $93.30$ & $88.14$ & $98.33$ & $88.76$ & $34.65$ \\ 
Neural Network & $93.54$ & $88.80$ & $96.52$ & $90.74$ & $48.31$ \\ 
XGBoost & $93.58$ & $88.84$ & $96.92$ & $90.47$ & $46.41$ \\ 
\midrule
\multicolumn{6}{l}{5-Min Median} \\ 
\midrule
Decision Tree & $95.48$ & $91.87$ & $98.02$ & $93.07$ & $48.01$ \\ 
Logistic Regression & $95.37$ & $91.55$ & $99.14$ & $91.87$ & $37.47$ \\ 
Neural Network & $95.75$ & $92.33$ & $98.39$ & $93.24$ & $49.14$ \\ 
XGBoost & $95.78$ & $92.41$ & $98.23$ & $93.45$ & $50.90$ \\ 
\midrule
\multicolumn{6}{l}{10-Min Median} \\ 
\midrule
Decision Tree & $96.02$ & $92.75$ & $98.30$ & $93.84$ & $47.95$ \\ 
Logistic Regression & $96.04$ & $92.73$ & $99.20$ & $93.09$ & $40.56$ \\ 
Neural Network & $96.18$ & $93.06$ & $98.29$ & $94.17$ & $50.90$ \\ 
XGBoost & $96.60$ & $93.79$ & $99.10$ & $94.22$ & $50.98$ \\ 
\bottomrule
\end{longtable}

Analyzing the area under the curve for precision-recall (AUC-PR) for all
models across raw ZM predictions and 5- and 10-minute median filtered
predictions yielded distinct results. Across all types of ZM
predictions, the decision tree consistently yielded the highest AUC-PR,
whereas the Neural Network generally performed least effectively. This
highlights the superior predictive accuracy of the decision tree model
in this context (see Figure~\ref{fig-pr_curves}). The Area Under the
Receiver Operating Characteristics Curve (ROC-AUC) results indicate a
varying performance of the four models across the ZM predictions types.
The XGBoost model consistently outperforms the other models across all
data types, suggesting its superior capacity for distinguishing between
classes. In contrast, the Neural Network model consistently exhibits the
lowest ROC-AUC values, suggesting a weaker performance in this specific
context (see Figure~\ref{fig-roc_curves}).

{\centering \includegraphics{visuals/plot_sleep_pr.pdf}

\caption{\label{fig-pr_curves}Precision-Recall curves of the models
evaluated across the different ZM predictions, including raw ZM
predictions, as well as 5-minute and 10-minute median smoothing of the
ZM raw predictions. The x-axis of the plot represents the proportion of
true wake epochs that were correctly classified as wake, while the
y-axis represents the proportion of all epochs labeled as wake by the
classifier that were classified correctly. The AUC values displayed in
the plot indicate the area under the Precision-Recall curve for each
model and condition.}

{\centering \includegraphics{visuals/plot_sleep_roc.pdf}

\caption{\label{fig-roc_curves}Text bla bla}
